### 2014-04-10 doc/uist14/abstract.tex
pass over abstract, intro, some work on related work

\bjoern{I prefer to write the abstract {\em first}, not last, as a blueprint for the rest of the paper.}

Emerging head-worn computing devices can enable hands-free interactions in physical spaces. 
%
We introduce and characterize a method for selecting and interacting with physical targets at a distance using head orientation. We augment a commercial wearable device, Google Glass, with an infrared (IR) emitter to select targets equipped with IR receivers. The fundamental imprecision of head orientation requires area selection techniques, which can lead to ambiguous selections. We introduce a two level disambiguation strategy: first, networked receivers use IR intensity sensors to estimate which target is the intended selection; second, users can fall back to a manual disambiguation technique using their near-eye display. The technique does not require an a priori map of target devices. However, a spatial data structure can be inferred to overcome sensor saturation problems in environments hostile to IR communication (e.g., in sunlight). \bjoern{speculation, if we get to it}
%
A target acquisition study with 14 participants shows that selection using head orientation with our device outperforms list selection on a wearable device. Intensity-based target disambiguation reduces selection time and error rate in dense target environments. \bjoern{rewrite once we have results}. We implement this technique in an application for controlling smart appliances and report qualitative user impressions.

>>>>

Emerging head-worn computing devices can enable hands-free interactions with smart objects in physical spaces. 
%
We introduce and characterize a method for selecting and interacting with physical targets at a distance using head orientation. We augment a commercial wearable device, Google Glass, with an infrared (IR) emitter to select targets equipped with IR receivers. The fundamental imprecision of head orientation requires area selection techniques, which can lead to ambiguous selections. We introduce a two level disambiguation strategy: first, networked receivers use IR intensity sensors to estimate which target is the intended selection; second, users can fall back to a manual disambiguation technique using their near-eye display. The technique does not require an a priori map of target devices. However, a spatial data structure can be inferred to aid ordering of items during manual disambiguation.
%
A target acquisition study with 14 participants shows that selection using head orientation with our device outperforms list selection on a wearable device. Intensity-based target disambiguation reduces selection time and error rate in dense target environments. \bjoern{rewrite once we have results}. We implement this technique in an application for controlling smart appliances and report qualitative user impressions.

### 2014-04-10 doc/uist14/relatedwork.tex
pass over abstract, intro, some work on related work

\section{Related Work}
\bjoern{TODO: Add discussion of ~\cite{swindells_that_2002,merrill_augmenting_2007}.}
\bjoern{Need a discussion of area cursors~\cite{kabbash1995prince,worden1997making}.}

>>>>

%!TEX root = uist14.tex
\section{Background and Related Work}
\subsection{Head and Eye Input}
Head movement has long been used for virtual camera control in VR applications~\cite{pausch_user_1993} and as an assitive input technology for cursor control of desktop applications~\cite{}. Head joysticks have been evaluated using Fitts' law and \bjoern{what did they find?} However, 
Card, summarizing other work, shows that neck muscles are a poor muscle group for pointing in general: neck muscles only have a bandwidth of about $4.2bits/s$ (compared to $23bits/s$ of wrist muscles used by a standard mouse)~\cite{Card:1991:MAD:123078.128726}.

Contrast: we're not working on GUI tasks, but on identifyign and selecting ubicomp objects in physical spaces.

Gaze control of existing graphical user intefaces~\cite{kumar2007eyepoint}. There are also wearable gaze trackers~\cite{SMI} -- but while these can provide a concrete point in space where a user is looking, turning this information into a selection information requires either a fixed absolute map of each space with known location and orientation of the user, or... Our system works through point-to-point IR communication and does not require any a priori knowledge of where objects are located.

Our work is closer in spirit to Selker's headworn system~\cite{Selker:2001:EGE:634067.634176}.

\subsection{Area Cursors}
A central hypothesis of this paper is that the area selection paradigm is well matched to head orientation input.
\bjoern{Need a discussion of area cursors~\cite{kabbash1995prince,worden1997making,Findlater-uist2010}.}

\subsection{Pointing in physical spaces}
\bjoern{take this from prior CHI submission.}

\bjoern{TODO: Add discussion of ~\cite{swindells_that_2002,merrill_augmenting_2007}.}
